---
title: "GlasshouseTrials_Pcontorta"
author: "Krisztian Nemeth"
date: "2024-03-05"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("ggpubr", quietly = TRUE))
  install.packages("ggpubr")
library(ggpubr)

if (!require("ggplot2", quietly = TRUE))
  install.packages("ggplot2")
library(ggplot2)

if (!require("ggstats", quietly = TRUE))
  install.packages("ggstats")
library(ggstats)

if (!require("tinytex", quietly = TRUE))
  install.packages("tinytex")
library(tinytex)

knitr::opts_knit$set(root.dir = "C:/Users/krinem/OneDrive - UKCEH")

```

## R Markdown

```{r}
GlasshouseTrial_Pcontorta <- read.delim("GlasshouseTrial_Pcontorta.txt")

GlasshouseTrial_Pcontorta$Adjusted.Height..mm. <- as.numeric(GlasshouseTrial_Pcontorta$Adjusted.Height..mm.)

GlasshouseTrial_Pcontorta$DBB..mm. <- as.numeric(GlasshouseTrial_Pcontorta$DBB..mm.)

GlasshouseTrial_Pcontorta$Type <- as.factor(GlasshouseTrial_Pcontorta$Type)

GlasshouseTrial_Pcontorta$Provenance <- as.factor(GlasshouseTrial_Pcontorta$Provenance)

GlasshouseTrial_Pcontorta$Collection.site..UK. <- as.factor(GlasshouseTrial_Pcontorta$Collection.site..UK.)


```


```{r , fig.cap = c("Figure 1:  ","Figure 2: ")}

my_comparisons <- list( c("Origin", "Plantation"), c("Origin", "Regeneration"), c("Plantation", "Regeneration") )


###Height 

TypePlot_Height <- ggplot(GlasshouseTrial_Pcontorta, aes(x=Type, y=Adjusted.Height..mm., fill=Type)) +
  geom_boxplot(alpha=0.7) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Set1") +
  xlab("Seed source") + ylab("Height (mm)") +
  stat_compare_means(method= "t.test", comparisons = my_comparisons, label.y = c(80, 90, 100), hide.ns = TRUE) +
  stat_compare_means(method= "anova", label.y = 100)
TypePlot_Height

##DBB
TypePlot_DBB <- ggplot(GlasshouseTrial_Pcontorta, aes(x=Type, y=DBB..mm., fill=Type)) +
  geom_boxplot(alpha=0.7) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Set1") +
  xlab("Seed source") + ylab("DBB (mm)") +
  stat_compare_means(method= "t.test", comparisons = my_comparisons, label.y = c(0.90, 1, 1.1), hide.ns = TRUE) +
  stat_compare_means(method= "anova", label.y = 1.1)
TypePlot_DBB

TypePlot_ProportionMortality <- ggplot(GlasshouseTrial_Pcontorta) +
  aes(x = Type, fill = Status, by = Type) +
  geom_bar(position = "fill", alpha=0.7) +
  geom_text(stat = "prop", position = position_fill(.5)) +
      xlab("Type") + ylab("Proportion") +
  scale_fill_brewer(palette = "Set1", name="Status")
TypePlot_ProportionMortality

TypePlot_CountMortality <- ggplot(GlasshouseTrial_Pcontorta) +
  aes(x = Type, fill = Status, by = Type) +
  geom_bar(alpha=0.7) +
  geom_text(stat = "prop", position = position_stack(.5)) +
      xlab("Type") + ylab("Count") +
  scale_fill_brewer(palette = "Set1", name="Status")
TypePlot_CountMortality


```
```{r, fig.width=10, fig.height=10, fig.fullwidth=TRUE, echo=FALSE,message=FALSE}

###Height 

OriginProvenancePlot_Height <- ggplot(GlasshouseTrial_Pcontorta, aes(x=Provenance, y=Adjusted.Height..mm., fill=Type)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_brewer(palette="Set1") +
  xlab("Provenance") + ylab("Height (mm)") +
  stat_compare_means(method= "anova", label.y = 90) 
OriginProvenancePlot_Height

##DBB
OriginProvenancePlot_DBB <- ggplot(GlasshouseTrial_Pcontorta, aes(x=Provenance, y=DBB..mm., fill=Type)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_brewer(palette="Set1") +
  xlab("Provenance") + ylab("DBB (mm)") +
  stat_compare_means(method= "anova", label.y = 1) 
OriginProvenancePlot_DBB

OriginProvenancePlot_ProportionMortality <- ggplot(GlasshouseTrial_Pcontorta) +
  aes(x = Provenance, fill = Status, by = Provenance) +
  geom_bar(position = "fill", alpha=0.7) +
  geom_text(stat = "prop", position = position_fill(.5)) +
      xlab("Provenance") + ylab("Proportion") +
  scale_fill_brewer(palette = "Set1", name="Status") +
  facet_grid(cols = vars(Type), scales = "free")
OriginProvenancePlot_ProportionMortality

OriginProvenancePlot_CountMortality <- ggplot(GlasshouseTrial_Pcontorta) +
  aes(x = Provenance, fill = Status, by = Type) +
  geom_bar(alpha=0.7) +
  geom_text(stat = "prop", position = position_stack(.5)) +
      xlab("Provenance") + ylab("Count") +
  scale_fill_brewer(palette = "Set1", name="Status") +
    facet_grid(cols = vars(Type), scales = "free")
OriginProvenancePlot_CountMortality

```


```{r}

###Height 

GlasshouseTrial_Pcontorta_UKCollections <- GlasshouseTrial_Pcontorta[which(GlasshouseTrial_Pcontorta$Type %in% c("Plantation", "Regeneration")), ]

CollectionSitePlot_Height <- ggplot(GlasshouseTrial_Pcontorta_UKCollections, aes(x=Collection.site..UK., y=Adjusted.Height..mm., fill=Collection.site..UK.)) +
  geom_boxplot(alpha=0.7) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Set1") +
  xlab("Collection Site (UK)") + ylab("Height (mm)") +
  stat_compare_means(method = "t.test",label.x = 1.5, label.y = 80) 
CollectionSitePlot_Height

##DBB

CollectionSitePlot_DBB <- ggplot(GlasshouseTrial_Pcontorta_UKCollections, aes(x=Collection.site..UK., y=DBB..mm., fill=Collection.site..UK.)) +
  geom_boxplot(alpha=0.7) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Set1") +
  xlab("Collection Site (UK)") + ylab("DBB (mm)") +
  stat_compare_means(method = "t.test",label.x = 1.5, label.y = 1) 
CollectionSitePlot_DBB

CollectionSitePlot_ProportionMortality <- ggplot(GlasshouseTrial_Pcontorta_UKCollections) +
  aes(x = Collection.site..UK., fill = Status, by = Collection.site..UK.) +
  geom_bar(position = "fill", alpha=0.7) +
  geom_text(stat = "prop", position = position_fill(.5)) +
      xlab("Collection site( )UK)") + ylab("Proportion") +
  scale_fill_brewer(palette = "Set1", name="Status")
CollectionSitePlot_ProportionMortality

CollectionSitePlot_CountMortality <- ggplot(GlasshouseTrial_Pcontorta_UKCollections) +
  aes(x = Collection.site..UK., fill = Status, by = Collection.site..UK.) +
  geom_bar(alpha=0.7) +
  geom_text(stat = "prop", position = position_stack(.5)) +
      xlab("Collection site (UK)") + ylab("Count") +
  scale_fill_brewer(palette = "Set1", name="Status")
CollectionSitePlot_CountMortality
```
 
 Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


