---
title: "National Trust Samples"
author: "Krisztian Nemeth"
date: "2024-03-11"
output:
  word_document: default
  html_document: default
  fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("plotly", quietly = TRUE))
  install.packages("plotly")
library(plotly)

if (!require("ggfortify", quietly = TRUE))
  install.packages("ggfortify")
library(ggfortify)

if (!require("data.table", quietly = TRUE))
  install.packages("data.table")
library(data.table)

if (!require("adegenet", quietly = TRUE))
  install.packages("adegenet")
library(adegenet)

if (!require("vcfR", quietly = TRUE))
  install.packages("vcfR")
library(vcfR)

if (!require("LEA", quietly = TRUE))
  install.packages("LEA")
library(LEA)

if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
library(BiocManager)

knitr::opts_knit$set(root.dir = "C:/Users/krinem/OneDrive - UKCEH/National Trust Samples")
```

## Introduction

16 Samples from two putative native trees collected at two National Trust Sites; Sheringham Estate (n = 13), Felbrigg Estate (n = 3) were genotyped and analysed alongside 194 authocthanous samples from sites across Scotland. Principal Component analysis was performed to detect any underlying population structure. Additional Admixture analysis was performed using the LEA software to identify the ancestry components. The samples were also analysed against samples from across Europe using the same methods. 

```{r Data Scotland}
# Read in TASSEL 5 PCA values in Plink format from a text file into a data frame
df <- read.table("PC_newLEAFgenotypes_Mothers+NationalTrustSamples.txt", header = TRUE, sep = "\t")

# Read a Keyfile containing Taxa names and corresponding information of interest into a data frame
Keyfile <- read.table("newLEAFgenotypes_Mothers+NationalTrustSamples_Keyfile.txt", header = TRUE, sep = "\t")

# Merge the two data frames based on specific columns
merged_data <- merge(df, Keyfile, by.x = "FID", by.y = "Taxa", row.names = FALSE)
merged_data$Classification <- factor(merged_data$Classification, levels=c("Cairngorms", "Central Atlantic", "Hyper-oceanic", "Sheringham Estate", "Felbrigg Estate"))

NationalTrust_data <- subset(merged_data, Classification %in% c("Sheringham Estate", "Felbrigg Estate"))

```

## Including Plots


```{r PC Scotland, echo=FALSE}
# Create a ggplot2 plot
ClassificationPC <- ggplot(merged_data, aes(x = PC1, y = PC2, color = Classification)) +
  geom_point() +
  labs(x = "Principal Component 1 (1.111%)",
       y = "Principal Component 2 (1.076%)") +
  scale_color_manual(values = c("#b69140",
                                 "#8d70c9",
                                 "#6aa74d",
                                 "#c8588c",
                                 "#49adad")) +
  stat_ellipse(data = NationalTrust_data, aes(group = NULL), color="red", type = "norm") +
  labs(color = "Classification")  # Change the legend title to "Classification"

# Convert the ggplot2 plot to an interactive Plotly plot
interactive_plotClassificationPC <- ggplotly(ClassificationPC, originalData = FALSE)  # Set originalData to FALSE

# Display the interactive plot
interactive_plotClassificationPC

SourcePC <- ggplot(merged_data, aes(x = PC1, y = PC2, color = Source)) +
  geom_point() +
  labs(x = "Principal Component 1 (1.111%)",
       y = "Principal Component 2 (1.076%)") +
  stat_ellipse(data = NationalTrust_data, aes(group = NULL), color="red", type= "norm") +
  labs(color = "Source")  # Change the legend title to "Source"

# Convert the ggplot2 plot to an interactive Plotly plot
interactive_plotSourcePC <- ggplotly(SourcePC, originalData = FALSE)  # Set originalData to FALSE

# Display the interactive plot
interactive_plotSourcePC

```


```{r DAPC Scotland, echo=FALSE}
dat.vcf <- read.vcfR("newLEAFgenotypes_Mothers+NationalTrustSamples.vcf")

dat.genind <- vcfR2genind(dat.vcf)
dat.genind        

Populations_data <- read.table("newLEAFgenotypes_Mothers+NationalTrustSamples_popfile.txt", header = TRUE, sep = "\t")

## Add population identifier
dat.genind$pop <- as.factor(Populations_data$pop)
cols <- hcl.colors(length(levels(pop(dat.genind))), palette="viridis", alpha=0.8, rev=F, fixup=T)

pop.info <- Populations_data

# we then run the first DACP to find out how many PCs we want to use in the final DACP (dacp4)
dapc3 <- dapc(dat.genind, n.pca=250, n.da=200, center=T, scale=F, NA.method="mean") # 150 PCs and 200 DFs retained

par(mfrow=c(1,1), oma=c(0.5,0.5,0.5,0.5), mar=c(3,3,3,3), family="Helvetica")
temp <- optim.a.score(dapc3, n.pca=1:ncol(dat.genind$tab)) # The number of retained PCs can be chosen so as to optimize the a-score; this is achived by optim.a.score
#

n_pca <- 177


par(mfrow=c(3,1))
for (clust in 2:10) 
{
  grp <- find.clusters(dat.genind, n.pca=150, n.clust=clust)
  dapc5 <- dapc(dat.genind, grp$grp, n.pca=n_pca, n.da=10, center=T, scale=F) 
  assign("ap", cbind(pop.info,dapc5$posterior))
  #for the comparison with sNMF, we also write a file with assignment probabilities
  write.table(ap, paste("ap_DAPC_K", clust,".txt", sep=""), sep="\t", quote=F, row.names=F)
  ind.nb <- as.data.frame(table(ap$pop)); colnames(ind.nb) <- c("Population","Individuals")
  axis.position <-c()
  assign("ap_1", as.matrix(t(ap)))
  barplot(ap_1[3:(3+clust-1),], main=paste("Posterior assignment probailities K=", clust, sep=""), col=viridis(clust), space=0,border=NA, xaxt='n', las=3)
  for (j in 1:NROW(ind.nb)) 
  {
    axis.position[j+1] <- sum(axis.position[j], ind.nb[j,"Individuals"])
    abline(v=axis.position, col="black", lty=1, lwd=1)
    text.position <- NULL
    text.position <- axis.position[j+1] - (ind.nb[j,"Individuals"] / 2)
    text(text.position, -0.2, ind.nb[j,"Population"], srt=90, cex=1.0, xpd=T, las=1)
  }   
}


```

```{r Data Europe}
# Read in TASSEL 5 PCA values in Plink format from a text file into a data frame
df <- read.table("PC_Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.txt", header = TRUE, sep = "\t")

# Read a Keyfile containing Taxa names and corresponding information of interest into a data frame
Keyfile <- read.table("Keyfile_Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.txt", header = TRUE, sep = "\t")

# Merge the two data frames based on specific columns
merged_data <- merge(df, Keyfile, by.x = "FID", by.y = "Taxa", row.names = FALSE)

merged_data$Source <- factor(merged_data$Source, levels=c("Ireland", "Scotland", "Sheringham Estate", "Felbrigg Estate", "Spain", "Italy", "Switzerland", "Austria", "Poland", "Bulgaria", "Sweden", "Finland", "Siberia", "Turkey"))

NationalTrust_data <- subset(merged_data, Source %in% c("Sheringham Estate", "Felbrigg Estate"))

```

```{r PC Europe, echo=FALSE}
# Create a ggplot2 plot

PC <- ggplot(merged_data, aes(x = PC1, y = PC2, color = Source)) +
  geom_point() +
  labs(x = "Principal Component 1 (3.774%)",
       y = "Principal Component 2 (2.534%)") +
  scale_color_manual(values = c("#bb486a",
"#c288d3",
"#8fb745",
"#6d71d8",
"#c49c38",
"#563686",
"#4cc490",
"#be62c1",
"#709243",
"#b14a89",
"#bf6f39",
"#5e8bd5",
"#b64741",
"#bb486a")) +
  stat_ellipse(data = NationalTrust_data, aes(group = NULL), color="red", type = "norm") +
  labs(color = "Source")  # Change the legend title to "Classification"

# Convert the ggplot2 plot to an interactive Plotly plot
interactive_plotPC <- ggplotly(PC, originalData = FALSE)  # Set originalData to FALSE

# Display the interactive plot
interactive_plotPC

```

```{r SNMF, echo=FALSE}
# Convert VCF file to LFMM format and save it
vcf2lfmm("Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.vcf", output.file = "Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples", force = TRUE)

# Perform Principal Component Analysis (PCA) on LFMM data
pc = pca("Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.lfmm", scale = TRUE)

# Calculate Tracy-Widom statistics
tw = tracy.widom(pc)

# Plot the percentage of variance explained by each component
plot(tw$percentage, pch = 19, col = "darkblue", cex = .8)

# Create an SNMF project
project = snmf("Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.geno",
               K = 1:12,
               entropy = TRUE,
               repetitions = 10,
               project = "new")

```

```{r Admixture, echo=FALSE}

project <- load.snmfProject("Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.snmfProject")

# Plot the cross-entropy criterion for all runs in the SNMF project
plot(project, col = "blue", pch = 19, cex = 1.2)

# Select the best run for K = 6 clusters
best = which.min(cross.entropy(project, K = 6))

# Read population data from a file
Populations_data <- read.table("geno_poplist_Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.txt", header = TRUE, sep = "\t")

# Read the desired order from a file
desired_order <- read.csv("desired_order_Europe+Scotland_FilteredGenotypes_Proportional_NationalTrustSamples.txt", header = FALSE)[, 1]

# Calculate the Q-matrix for the best run
qmatrix <- LEA::Q(project, K = 6, run = best)
qmatrix <- as.data.frame(qmatrix)

# Combine population data and Q-matrix, and sort by desired order
tbl <- cbind(Populations_data, qmatrix)
tbl <- tbl[order(match(tbl$taxa, desired_order)), ]

# Define colors for the barplot
col = c("#50ac72",
        "#c45ca2",
        "#929d3d",
        "#7879cd",
        "#c8803e",
        "#cc5452")

# Find breakpoints between different populations
breaks <- 0
for (i in 2:nrow(tbl)) {
  if (tbl$pop[i] != tbl$pop[i - 1]) {
    breaks <- c(breaks, i - 1)
  }
}
breaks <- c(breaks, nrow(tbl))

# Calculate spaces and labels for the barplot
spaces <- rep(0, nrow(tbl))
spaces[breaks[2:(length(breaks) - 1)] + 1] <- 2
labels <- rep("", nrow(tbl))
labels[breaks[1:(length(breaks) - 1)] + ((breaks[2:length(breaks)] - breaks[1:(length(breaks) - 1)]) / 2)] <- gsub("_", " ", tbl$pop[breaks])

# Set plot parameters
par(oma = c(4.5, 1, 1, 1), mar = c(3, 3, 2, 1), mgp = c(2, 1, 0), xpd = TRUE)

# Create the barplot
barplot(t(as.matrix(tbl[, 3:8])), col = col, xaxt = "n", border = NA, xlim = c(0, nrow(tbl)), width = 1, space = spaces, names.arg = rep("", nrow(tbl)), ylab="Ancestry proportions", main=paste("K = ",opt.K,sep=""), cex.lab = 1)

# Add x-axis labels with line breaks
axis(1, line = -0.2, at = breaks[1:(length(breaks) - 1)] + ((0:(length(breaks) - 2)) * 2) + ((breaks[2:length(breaks)] - breaks[1:(length(breaks) - 1)]) / 2), lwd = 0, lwd.ticks = 1, labels = gsub("_", " ", tbl$pop[breaks]), las = 2, cex.axis = 0.8)

```

```{r DAPC, echo=FALSE}
dat.genind <- vcfR2genind(dat.vcf)
dat.genind        
## Add population identifier
dat.genind$pop <- as.factor(Keyfile$Classification)
cols <- hcl.colors(length(levels(pop(dat.genind))), palette="viridis", alpha=0.8, rev=F, fixup=T)

## Un-supervised version -> we determine the number of clusters by K-means
par(mfrow=c(1,1))
kval <- find.clusters(dat.genind, n.pca=150, max.n.clust=60, choose.n.clust=F)
#wait for it!

## take K=8 for FS and K=3 for PS
k <- 2
# we run DAPC only with the K determined above. For that we need to know the cluster membership of each individual (grp$grp)
grp <- find.clusters(dat.genind, n.pca=150, n.clust=k, choose.n.clust=T)

# we then run the first DACP to find out how many PCs we want to use in the final DACP (dacp4)
dapc3 <- dapc(dat.genind, n.pca=250, n.da=200, center=T, scale=F, NA.method="mean") # 150 PCs and 200 DFs retained

# The trade-off between power of discrimination and over-fitting can be measured by the a-score, which is simply the difference between the proportion of successful reassignment of the analysis (observed discrimination) and values obtained using random groups (random discrimination)
cols <- hcl.colors(length(unique(grp$grp)), palette="viridis", alpha=0.8, rev=F, fixup=T)
#pdf(paste(dir.path1,"dapc/",codesp[sp],"/DAPC_OptimalScores_Kmeans.pdf",sep=""), width=8.27*1.0, height=8.27*0.7) # Represent both the genotypes and inertia ellipses for populations
par(mfrow=c(1,1), oma=c(0.5,0.5,0.5,0.5), mar=c(3,3,3,3), family="Helvetica")
temp <- optim.a.score(dapc3, n.pca=1:ncol(dat.genind$tab)) # The number of retained PCs can be chosen so as to optimize the a-score; this is achived by optim.a.score
#dev.off()
#choose the number of PCAs to be used
n_pca <- 30

#perform the final DAPC with the K defined above (dapc4)
dapc4 <- dapc(dat.genind, grp$grp, n.pca=n_pca, n.da=10, center=T, scale=F) 
# Calculate proportion of explained variance of the first two dimensions
(dapc4$eig[1]/sum(dapc4$eig))*100 
(dapc4$eig[2]/sum(dapc4$eig))*100
(sum(dapc4$eig[1:2])/sum(dapc4$eig))*100

# plot the DAPC showing clusters
#pdf(paste(dir.path1,"dapc/",codesp[sp],"/DAPC_Kmeans.pdf",sep=""), width=8.27*0.7, height=9.5) # Represent both the genotypes and inertia ellipses for populations
layout(matrix(c(1,1,2,3), 2, 2, byrow=T), widths=c(1,1), heights=c(1,0.5))
par(oma=c(0.5,0.5,0.5,0.5), mar=c(4,4,2,2), family="Helvetica")
scatter(dapc4, xax=1, yax=2, sub="DAPC 1-2", possub="topright", posi.da="bottomright", 
        clabel=1.2, bg="white", cstar=1, col=cols, solid=0.9, scree.pca=T, posi.pca="bottomleft")
text(2, 460, labels="a", cex=2, xpd=NA)
scatter(dapc4, 1, 1, col=cols, bg="white", scree.da=F, legend=F, solid=0.3)
text(-9.5, 0.46, labels="b", cex=2, xpd=NA)
scatter(dapc4, 2, col=cols, bg="white", scree.da=F, legend=F, solid=0.3)
text(-8.2, 0.48, labels="c", cex=2, xpd=NA)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
